### 2019-12-10 Statistical Methods/02_numerical-summaries.rmd
statistical methods, ch2 wrap-up
statistical methods, ch2 wrap-up

>>>>

### 2.29

The largest observation in the last problem is 49. If 49 is replaced by the value 200, verify that the sample mean is now $\bar{X} = 50.1$ but the 20% trimmed mean and median are not changed.

```{r, echo = T}
x <- c(21, 36, 42, 24, 25, 36, 35, 200, 32)

M <- o[ (n + 1)/2 ]
```

### 2.30

For the data in Exercise 28, what is the minimum number of observations that must be altered so that the 20% trimmed mean is greather than 1,000?

$g + 1 = `r g + 1`$

### 2.31

Repeat the previous exercise, but use the median instead.

_N = 4_

What does this illustrate about the resistance of the mean, median and 20% trimmed mean?

_The mean has the least resistance, followed by the trimmed mean and then the median._

### 2.32

6, 3, 2, 7, 6, 5, 8, 9, 11

Use R to verify that the sample mean, 20% trimmed mean, and median are $\bar{X} = 6.5, \bar{X}_t = 6.7 \ and \ M = 6.5$, respectfully.

```{r, echo = T}
x <- c(6, 3, 2, 7, 6, 5, 8, 9, 8, 11)

mean(x)

mean(x, trim = .2)

median(x)
```

### 2.33

In general, when you have _n_ observations, what is the minimum number of values that must be altered to make the 20% trimmed mean grow as larage as you want?

floor(n*.2) + 1

### 2.34

A class of fourth graders was asked to bring a pumpkin to school. Each of the 29 students counted the number of seeds in their pumpkin, and the results were:

250, 220, 281, 247, 230, 209, 240, 160, 370, 274, 210, 204, 243, 251, 190,
200, 130, 150, 177, 475, 221, 350, 224, 163, 272, 236, 200, 171, 98

Use R to compute the sample mean, 20% trimmed mean, median and MOM.

```{r, echo = T}
x <- c(250, 220, 281, 247, 230, 209, 240, 160, 370, 274, 210, 204, 243, 251, 190, 
       200, 130, 150, 177, 475, 221, 350, 224, 163, 272, 236, 200, 171, 98)

xbar <- mean(x)
tbar <- mean(x, trim = .2)
M <- median(x)
MOM <-  mean( x[ abs( x - median(x) ) / mad(x) < 2.27 ] )

$\bar{X} = `r xbar`$, $\bar{X}_t = `r tbar`$, $M = `r M`$, $MOM = `r MOM`$

### 2.35

Compute the 20% Winsorized values for the observations:

```{r, echo = T}
x <- c(21, 36, 42, 24, 25, 36, 35, 49, 32)

winsorize <- function( x, trim = .2) {
   n <- length(x)

   o <- sort(x)
   g <- floor(trim*n)

   o[1:(g+1)] <- o[(g+1)]
   o[(n-g):n] <- o[n-g]

   o
}

winsorize(x)
```

### 2.36

For the observations in the pervious problem, use R to verify that the 20% Winsorized variance is 51.36.

Wvar <- var(winsorize(x))

$W_{var} = `r Wvar`$

### 2.37

In the previous problem, would you expect the sample variance to be larger or smaller than 51.36?

_Larger, Winsorizing pulls in the extremes._

Verify your answer.

```{r, echo = T}
var(x)
```

### 2.38

In general, will the Winsorized sample variance, $s^2_w$, be less than the sample variance, $s^2$?

### 2.39

6, 3, 2, 7, 6, 5, 8, 9, 8, 11

verify that the sample variance and 20% Winsorized variance are 7.4 and 1.8, respectfully.

```{r, echo = T}
x <- c(6, 3, 2, 7, 6, 5, 8, 9, 8, 11)

variance <- var(x)

$Var = `r variance`$, $Var_w = `r Wvar`$

### 2.40

Consider again the number of pumpkin seeds given in Exercise 34.

Compute the 20% Winsorized variance.

```{r, echo = T}
x <- c(250, 220, 281, 247, 230, 209, 240, 160, 370, 274, 210, 204, 243, 251, 190, 
       200, 130, 150, 177, 475, 221, 350, 224, 163, 272, 236, 200, 171, 98)

Wvar <- var( winsorize(x) )

$W_{var} = `r Wvar`$

### 2.41

Snedecor and Cochran (1967) report results from an experiment dealing with weight gaim in rats as a function of source and amount of protein.

One of the groups was fed beef with a low amount of protein. The weight gains were:

90, 76, 90, 64, 86, 51, 72, 90, 95, 78

Compute the 20% trimmed mean and 20% Winsorized variance.

```{r, echo = T}
x <- c(90, 76, 90, 64, 86, 51, 72, 90, 95, 78)

tbar <- mean(x, trim = .2)
wvar <- var(winsorize(x))

$\bar{X}_t = `r tbar`$, $Var_w = `r wvar`$
